---
title: "Bioinformatics FYE"
author: "Madeleine Duquette (A59019542)"
format:
  pdf:
    geometry: margin=.25in
---

## COVID-19 Variants Plot

#### Install and load ggplot2, lubridate, and dplyr. Will also need to load readr.

```         
# install.packages("ggplot2")
# install.packages("lubridate")
# install.packages("dplyr")
```

```{r}
library(ggplot2)
library(lubridate)
library(dplyr)
library(readr)
```

#### Read in the csv file

```{r}
variants <- read_csv("covid19_variants.csv")
```

#### Check out what's in there

```{r}
head(variants)
```

#### I don't need columns 7 and 8 because they do not contain data so I will create a dataframe and then subset columns 1-6

```{r}
df <- data.frame(variants)
df_subset <- df[, -c(7,8)]
head(df_subset)
```

#### Now let's try graphing something!

```{r}
ggplot(data=df_subset) + 
  aes(x=date, y=percentage) + 
  geom_line()
```

#### Not much to see there... Let's try changing some colors by the variant strain type

```{r}
ggplot(data=df_subset) + 
  aes(x=date, y=percentage, col=variant_name) + 
  geom_line()
```

#### Much better! Let's add labels.

```{r}
ggplot(data=df_subset) + 
  aes(x=date, y=percentage, col=variant_name) + 
  geom_line() + 
  labs(y="Percentage of sequenced specimens", color="Variant Name") + 
  xlab("")
```

#### We will now add some granularity to the x-axis by scaling by month instead of by year.

```{r}
ggplot(data=df_subset) + 
  aes(x=date, y=percentage, col=variant_name) + geom_line() + 
  labs(y="Percentage of sequenced specimens", color="Variant Name") + 
  xlab("") + 
  scale_x_date(date_labels="%b %Y", date_breaks="1 month")
```

#### Well, that's definitely not readable. Let's try angling the labels!

```{r}
ggplot(data=df_subset) + 
  aes(x=date, y=percentage, col=variant_name) + 
  geom_line() + 
  labs(y="Percentage of sequenced specimens", color="Variant Name") + 
  xlab("") + 
  scale_x_date(date_labels="%b %Y", date_breaks="1 month") + 
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

# Yay! 
